---
title: "Appendix 3: Mechanisms behind functionally distinct species’ effect on productivity."
author:
  - name: "Delalandre et al."
abstract: >
  Functionally distinct tree species support long-term productivity in extreme environments
date: "`r format(Sys.time(), '%d %B %Y')`"
url: https://github.com/LDelalandre/Project-1_Distinct-sp_BEF
github-repo: GitHubID/Repository
lang: en-US
otherlangs: [fr-FR,it]
bibliography: BEF.bib
biblio-style: chicago
pdftoc: no
preamble: >
  \hyphenation{bio-di-ver-si-ty sap-lings}
always_allow_html: yes
output:
  bookdown::pdf_book:
    template: latex/template.tex
    citation_package: natbib
    latex_engine: xelatex
    keep_tex: yes
  rmdformats::downcute:
    use_bookdown: yes
    lightbox: yes
  bookdown::word_document2: default
  bookdown::gitbook:
    config:
      download: "pdf"
      sharing:
        github: yes
  bookdown::html_document2:
    toc: yes
    toc_float: yes
---

```{r DoNotModify, include=FALSE}
### Utilities. Do not modify.
# Installation of packages if necessary
InstallPackages <- function(Packages) {
  InstallPackage <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {
      install.packages(Package, repos="https://cran.rstudio.com/")
    }
  }
  invisible(sapply(Packages, InstallPackage))
}

# Basic packages
InstallPackages(c("bookdown", "formatR", "kableExtra", "ragg"))

# kableExtra must be loaded 
if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx") {
  # Word output (https://stackoverflow.com/questions/35144130/in-knitr-how-can-i-test-for-if-the-output-will-be-pdf-or-word)
  # Do not use autoformat (https://github.com/haozhu233/kableExtra/issues/308)
  options(kableExtra.auto_format = FALSE)
}
library("kableExtra")

# Chunk font size hook: allows size='small' or any valid Latex font size in chunk options
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r Options, include=FALSE}
### Customized options for this document
# Add necessary packages here
Packages <- c("tidyverse")
# Install them
InstallPackages(Packages)

# knitr options
knitr::opts_chunk$set(
  cache=FALSE, # Cache chunk results
  echo = TRUE, # Show/Hide R chunks
  warning=FALSE, # Show/Hide warnings
  # Figure alignment and size
  fig.align='center', out.width='80%',
  # Graphic devices (ragg_png is better than standard png)
  dev = c("ragg_png", "pdf"),
  # Code chunk format
  tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=50),
  size="scriptsize", knitr.graphics.auto_pdf = TRUE
  )
options(width=50)

# ggplot style
library("tidyverse")
theme_set(theme_bw())
theme_update(panel.background=element_rect(fill="transparent", colour=NA),
             plot.background=element_rect(fill="transparent", colour=NA))
knitr::opts_chunk$set(dev.args=list(bg="transparent"))

# Random seed
set.seed(973)
```

# Context

Manipulative studies usually report a diversity effect as an increase of the magnitude of an ecosystem function (for instance, productivity) in a polyculture relative to monocultures. A partitioning analysis of diversity effects  (Loreau & Hector, 2001) has been extensively used to separate it into selection (the fraction of this effect that is due to the presence of productive species in the mixture) and complementarity (the fraction due to an increase in species productivity when they are mixed) effects.

## Math

Equations in LaTeX format can be inserted in line, like $A=\pi r^2$ or isolated like $$e^{i \pi} = -1$$.

They can be numbered, see equation \@ref(eq:disk), using the `equation` environment:

\begin{equation}
  A = \pi r^2.
  (\#eq:disk)
\end{equation}


## Cross-references

Figures and tables have an automatically generated label, identical to the name of the code snippet prefixed with `fig:` and `tab:`.

For equations, the label is added manually by the code `(\#eq:xxx)` before the end of the equation.

Sections can be tagged by ending their title with `{#yyy}`.

Bookmarks can also be placed freely in the text with the command `(ref:zzz)`.

In all cases, the call to the reference is made by the command `\@ref(ref:zzz)`.


## Bibliography

Bibliographic references included in the `references.bib` file can be called by `[@CitationKey]`, in parentheses [@Xie2016], or without square brackets, in the text, as @Xie2018 .


The bibliography is processed by Pandoc when producing Word or HTML documents.
The bibliographic style can be specified, by adding the line 

    csl:file_name.csl
    
in the document header and copying the _.csl_ style file to the project folder.
More than a thousand styles are available[^4].

[^4]: <https://github.com/citation-style-language/styles>

For PDF documents, the bibliography is managed by natbib.
The style is declared in the header:

    biblio-style: chicago

It can be changed as long as the appropriate `.bst` file (by default: `chicago.bst`) is included in the project. 


## LaTeX preamble

LaTeX commands can be added in the preamble of the produced LaTeX file, for example to load additional packages.
These commands are in the `preamble:` section of the Markdown file header.

The default commands allow to show the use of the hyphenation command:
```
\hyphenation%
  {bio-di-ver-si-ty sap-lings}
```

Other commands can be added as needed.
Warning:

- Comments are not allowed
- Complex commands (such as `\renewenvironment`) must be entered on a single line otherwise they will be destroyed by knitr at the first knitting in HTML.


## Forcing line breaks

Hyphenation is handled automatically in LaTeX.
If a word is not hyphenated correctly, add its hyphenation in the preamble of the file with the command `hyphenation` (words are separated by spaces, hyphenation locations are represented by dashes).

If LaTeX can't find a solution for the line break, for example because some code is too long a non-breaking block, add the LaTeX command `\break` to the line break location.
Do not leave a space before the command. 
The HTML document ignores LaTeX commands.


## Languages

Languages are declared in the document header.

The main language of the document (`lang`) changes the name of some elements, such as the table of contents.
The change of language in the document (one of `otherlangs`) is managed in LaTeX but not in HTML by inserting on a new line the following command: 
```
\selectlanguage{english}
```

The current language has an effect only in LaTeX output: a space is added before double punctuation in French, the size of spaces is larger at the beginning of sentences in English, etc.
The `\selectlanguage` command is simply ignored in HTML.

Language codes are used in the header, such as `en-US` but language names are necessary in `\selectlanguage{}`.
Name matches are listed in table 3 of the polyglossia package documentation[^5].

[^5]: http://mirrors.ctan.org/macros/unicodetex/latex/polyglossia/polyglossia.pdf


# Document types

This template is designed to work with the Simple Article template in LaTeX and produce documents in PDF, HTML or Word format.
Use the list of choices in the _Knit_ button to choose the output format.

## PDF Document

The document is formatted according to the article LaTeX template.

## HTML document

The GitBook template is optimized for on-screen reading.
While writing, prefer knitting to HTML format for its speed of execution.
A download button is available in the document menu bar: it will work if the document is also knitted in PDF format and if the file name is entered in the download field of the YAML header.

The HMTL Document and all formats from the **rmdformats** packages are other possibilities.

## Word document

Its content can be formatted or copied into a template. 
The standard text styles are "First Paragraph" and "Body Text".

The advantage of the Word format is to produce a manuscript for journals that do not support LaTeX.
The bibliographic style of the journal is most likely available in _.csl_ format, which minimizes manual preparation.

The last line of this template (R code snippet) must be kept to display the title _References_ (to be translated into the document language if necessary) in HTML format.
The level 1 title _References_ must be added manually to Word files.

`r if (!knitr:::is_latex_output()) '# References {-}'`
